Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-07-30 paper.tex
Update paper.tex

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are commonly used to probe electron-vibration coupling in isotropic media.
Most of these methods rely on infrared and/or Raman transitions but methods involving hyper-Raman transitions are also possible. 
Nonlinear spectroscopy involving hyper-Raman transitions provides a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the structure and dynamics of isotropic systems.
Hyper Difference Frequency Generation (HDFG) spectroscopy is an underdeveloped four wave mixing vibrational spectroscopy based on the hyper-Raman transition. 
Unlike other coherent four-wave mixing techniques, HDFG only relies on coupling between a vibrational and electronic state.
Despite several experimental reports on HDFG, its spectroscopic properties have not been fully explored.  
To this end, we investigate the selection rules and behavior of HDFG spectroscopy as an upconverted infrared spectroscopy and as a probe of vibronic coupling in molecular systems.
HDFG is found to possess selection rules based upon infrared absorption spectroscopy and hyper-Raman scattering, making it sensitive to vibronic coupling mechanisms in isotropic systems. 
Applying the selection rules found for vibronic hyper-Raman scattering, it is found that HDFG output is dependent upon potential well offset.
It is found that HDFG and vibrational sum frequency generation (vSFG) spectroscopy share similar output, making HDFG feasible for vSFG practitioners. 
\end{abstract}

@@

\subsection{Electronically resonant and pre-resonant: Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to those found in Albrecht's treatment of Raman spectroscopy.\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(Q)) \otimes \ket{b}$ for electronic states $\{|a(Q))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where $Q$ is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on $Q$ for simplicity.

@@

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$ are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev', g0}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | v' \rangle \mel{v'}{Q_z}{0} \frac{1}{\Delta_{ev', g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev', g0}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where $\langle a | b \rangle$ and $\mel{a}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively. 
Duschinsky rotation effects are ignored in this discussion. \cite{Duschinsky1937}
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

For the case of electronic resonance, further simplifications can be made.
We restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite quanta in ground state vibrational modes. 
By working only in terms of a single normal mode and assuming the infrared pulse selectively excites $\ket{g,0} \rightarrow \ket{g,1}$ (i.e., removing the sum over $z$ and taking $v=1$), 
the A and B terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{v'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | v' \rangle
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar} \sum_{v'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | v' \rangle \mel{v'}{Q}{0} 
				\frac{1}{\Delta_{ev', g0}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{v'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{v'} 
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
			\end{split}
		\end{equation}
	\end{subequations}
where $\Lambda^{eg}_{0,jk}$ is the electronic two photon transition moment between $g$ and $e$.\cite{McClain1977}
$\Lambda$ contains the summation over $m$ in \autoref{ABterms} (N.B., here we use $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $).
Here, we restrict $v' = \{0,1,2\}$ for simplicity.

The impact of vibronic coupling in the HDFG spectrum is made clear by using a simple model. \cite{Kundu2022}
Using a two-well system described by 
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\xi)^2 \right) + \hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ (momentum) and $q$ (position) are the dimensionless conjugate coordinates of the normal mode, the A and B contributions can be evaluated using Franck-Condon and Herzberg-Teller integrals tabulated in terms of the dimensionless normal mode $q$ and offset parameter $\xi$ (\autoref{fig:doubres_spec}a). \cite{Carlson1988thesis} 
Note that $q = \sqrt{\frac{m^2\omega}{\hbar}} Q$.
The spectrum is simulated for an electronic state which is both one and two photon allowed.
A simulated spectrum (\autoref{fig:doubres_spec}b) highlighting the contributions of these terms on potential wells with $\xi = 0.5$ dissects the $A$ and $B$ contributions to the HDFG spectrum.
This stimulated hyper-Raman type spectrum would be the expected result in a HDFG experiment when investigating the vibronic structure of the electronic state coupled to $\ket{g,1}$.

@@

\autoref{ABterms_DR} shows important selection rules for the vibronic transitions.
We stress that while the $A,B$ selection rules are based upon any hyper-Raman active mode, the $\mel{v}{\mu_l}{0}$ term in \autoref{gammaabc} imposes the constraint that the $A,B$ transitions occur for infrared active normal mode.
The $A$ term is allowed whenever an electronic transition has a one- and two- photon transition moment.
The brightness is controlled by vibrational overlap (Franck-Condon factors).
Note that in the limit of complete potential energy surface overlap ($\xi \rightarrow 0$), $A$ vanishes as $\lim_{\xi \rightarrow 0} \langle 1 | 0 \rangle = 0$ for harmonic potential wells.\cite{Myers1982}
This selection rule is imposed by the dipole transition in HDFG. %todo: note this is the OPPOSITE of hyper raman
$B$ does not vanish in the same limit, as $\lim_{\xi \rightarrow 0} \mel{0}{Q}{1} = 1/\sqrt{2}$ for wells with identical curvature.
For centrosymmetric molecules, an electronic transition cannot be both one- and two- photon allowed by parity, so the $A$ term will vanish.
This makes HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects.\cite{Olson2018}
In the absence of an $A$ term, the $B$ terms will dominate.
The $B_1$ and $B_2$ terms involve different types of vibronic coupling pathways.
In $B_1$, the Herzberg-Teller coupling is associated with the two-photon absorption process, whereas in $B_2$, the Herzberg-Teller coupling is associated with one-photon emission.

Simulating the HDFG spectrum for the simple two-well system allows dissection of selection rules and spectral signatures without the need for complex couplings.
The spectrum selectively probes coupling between the vibronic states and $\ket{g,0}$, $\ket{g,1}$.
Previous reports suggest that $A$ term contributions may become negligible when significantly detuned from a resonance such that $\Delta_{ev', g0} \approx \Delta_{eg}$, i.e., the resonance denominator loses $\ket{v'}$ dependence. \cite{Neddersen1989}
The lack of $\ket{v'}$ dependence in the resonance denominator causes $A$ to vanish through closure.
However, the simulated spectrum (\autoref{fig:doubres_spec}b) shows a non-negligible $A$ contribution for $\omega_2 < 20000$ cm$^{-1}$, the limit for this model system where $\abs{\omega_{ev',g0} - 2\omega_2} \gg 2\Gamma_{ev',g0}$.
In other words, it appears difficult to escape the pre-resonance regime of hyper-Raman scattering.
Similar ideas have been seen in spontaneous Raman scattering. \cite{Warshel1977, Li1990, Gong2015}
As such, it may prove difficult to examine non-Condon effects ($B$ term contributions) in HDFG spectra of noncentrosymmetric systems.

In the resonant region, the $A$ and $B$ terms contain character from all three possible transitions between $\ket{g,0}$, $\ket{e,v'}$ and $\ket{g,1}$. 
This discussion assumes that mode symmetry allows for coupling between all of the vibronic states considered here.
Significant contributions from $\ket{e,0}$ and $\ket{e,1}$ in the $A$ term spectrum correspond well with expectations given the relative wavefunction overlap in \autoref{fig:doubres_spec}a, tabulated in \autoref{T:contributions}. 
Similarly, the $B_1$ term is dominated by coupling with the $\ket{e,1}$ state and $B_2$ is dominated by coupling with the $\ket{e,0}$ state (\autoref{T:contributions}). 
Note that contributions involving $\ket{e,2}$ are suppressed relative to the $\ket{e,0}$ and $\ket{e,1}$ contributions in the $A$ and $B$ terms.
This is expected, considering the large change in quantum number when involving $\ket{e,2}$ and the offset of the potential wells in \autoref{Hamiltonian}.

\begin{table}[!htbp]
	\caption{\label{T:contributions} Franck-Condon and/or Herzberg-Teller contributions to $A$, $B_1$, $B_2$ terms between $\ket{g,0}$, $\ket{g,1}$ and vibronic states $\ket{e,0}$, $\ket{e,1}$, $\ket{e,2}$ for \autoref{Hamiltonian} with an offset $\xi = 0.5$.
	Closed-form expressions for the Franck-Condon and Herzberg-Teller integrals used here have been tabulated by Carlson.\cite{Carlson1988thesis}}
	\begin{ruledtabular}
		\begin{tabular}{cccc}
			Resonance Term & $\ket{e,0}$ & $\ket{e,1}$ & $\ket{e,2}$\\
			\hline  
			$A$ 
			& $\langle 1 | 0 \rangle \langle 0 | 0 \rangle = 0.31 $  
			& $\langle 1 | 1 \rangle \langle 1 | 0 \rangle = 0.27 $ 
			& $\langle 1 | 2 \rangle \langle 2 | 0 \rangle = -0.04$\\

@@

\section{Quantitative Aspects of HDFG}\label{quant}
Despite their increasing use for dissecting intra and intermolecular interactions, four wave mixing techniques are usually not feasible for most laboratories, as they usually demand the use of multiple optical parametric amplifiers and/or complex acousto-optic modulators. \cite{Chen2016}
Unlike most four wave mixing techniques, HDFG can be performed using only two input beams (i.e., two color HDFG).
In the case of two color HDFG, only one tunable infrared optical parametric amplifier is needed, as the oscillator which pumps the infrared OPA can be used to perform the two-photon upconversion needed to generate HDFG output.
Vibrational sum frequency generation (vSFG) is a $\chi^{(2)}$ technique whose output, when only vibrationally resonant, scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor, whose phase-matching constraint is  $\vec{k}_3 = \vec{k}_1 + \vec{k}_2$.
HDFG is phase-matchable under $\vec{k}_4 = -\vec{k}_1 + 2\vec{k}_2$ conditions for most, if not all, vibrational resonances.
Laboratories which use vSFG to investigate interfacial species at buried interfaces thus have the ingredients necessary to perform a HDFG experiment in a transmission geometry. \cite{Piontek2023_1}
This would allow vSFG practitioners to perform measurements in the bulk, such as measuring steady-state vibrational spectra, free induction decay and vibrational population lifetimes through HDFG or pump-HDFG-probe.

To demonstrate the feasibility of HDFG for vSFG practitioners, we compare the relative output of each process. 
Macroscopically, oddness in spatial inversion of the vSFG polarization eliminates output from centrosymmetric species under the electric dipole approximation.\cite{RN132, RN133}
As a result, vSFG output is significantly reduced relative to most third order spectroscopies because it depends on surface number density, many orders of magnitude smaller than the bulk surface density, which is weak even for even for non-linear cross-sections.
However, HDFG depends upon hyper-Raman cross-sections, which are commonly several orders of magnitude weaker than corresponding Raman transitions.\cite{RN515}
As calculated in \autoref{appendixA}, these effects cancel, and the output polarization of HDFG and vSFG are roughly identical ($\abs{P_\text{vSFG} / P_\text{HDFG}} \sim 1$).
Transient HDFG, or IR-pump-HDFG-probe, analogous to IR-pump-vSFG-probe, is therefore also feasible under similar experimental conditions, and provides vSFG practitioners a simple method to measure bulk dynamics without significant changes in an experimental setup. 
Therefore, HDFG is a feasible spectroscopy for practitioners of vSFG both in terms of experimental setup and detection feasibility.

With the selection rules of HDFG understood for vibrational spectroscopy, and with HDFG generating a large enough output polarization, it becomes possible to obtain quantitative information from its spectra.
Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re(\chi^{(3)})$ and $\Im(\chi^{(3)})$.\cite{Levenson1974_1, Levenson1974_2}
In the method introduced by Levenson and Bloembergen (Bloembergen interferometry experiment), an internal standard interferes with the resonant lineshape.
Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
Resonant lineshapes are also complicated by amplitude level interference between the sample,  substrate and/or sample cell windows, which must be accounted for to obtain quantitatively correct $\chi^{(3)}$ values. \cite{RN362, RN418}
Most quantitative methods in an $n^{th}$ order CMDS experiment are used to measure $\chi^{(n)}$ values to compare the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.

Compared to $\vec{\mu}$ and $\alpha_{ij}$, it is difficult to calculate $\beta_{ijk}$ values from spontaneous hyper-Raman experiments. \cite{Kelley2010}
Only a few experimental determinations of $\beta_{ijk}$ for vibrational modes have been performed. \cite{Xu1997, Shoute2005, Kelley2010}
Methods reported in the literature to measure absolute hyper-Raman polarizabilities depend upon external standards such as hyperpolarizabilities of dissolved samples or two-photon absorption cross sections, also difficult to measure. \cite{Okuno2020}
Since the Bloembergen interferometry experiment only relies on the third order susceptibility for measured species (e.g., benzene or CaF$_2$),\cite{Levenson1974_2} it is possible to use quantitative four wave mixing spectroscopy to calculate $\beta_{ijk}$ values.
It is thus useful to investigate how a treatment of orientational averaging can extract $\beta_{ijk}$ from the HDFG $\chi^{(3)}_{IJKL}$ expression.

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are commonly used to probe electron-vibration coupling in isotropic media.
Most of these methods rely on infrared and/or Raman transitions, but methods involving hyper-Raman transitions are also possible. 
Hyper Difference Frequency Generation (HDFG) spectroscopy is an underdeveloped four wave mixing vibrational spectroscopy based on the hyper-Raman transition and IR absorption. 
HDFG is experimentally similar to vibrational sum frequency generation (vSFG) spectroscopy share similar output, making HDFG feasible for vSFG practitioners. 
Despite several experimental reports on HDFG, its spectroscopic properties have not been fully explored. 
To this end, we investigate the selection rules and behavior of HDFG spectroscopy as an upconverted infrared spectroscopy and as a probe of vibronic coupling in molecular systems.
HDFG is shown to a sensitive probe of vibronic coupling in bulk systems.
\end{abstract}

@@

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to those found in Albrecht's treatment of Raman spectroscopy.\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(Q)) \otimes \ket{b}$ for electronic states $\{|a(Q))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where $Q$ is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on $Q$ for simplicity.

@@

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$ are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev', g0}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | v' \rangle \mel{v'}{Q_z}{0} \frac{1}{\Delta_{ev', g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev', g0}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where $\langle a | b \rangle$ and $\mel{a}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively. 
Duschinsky rotation effects are ignored in this discussion. \cite{Duschinsky1937}
At this point, the expressions are valid for both electronically resonant and non-resonant cases.
For the case of electronic resonance, only a few components of this summation will be relevant ($e,v^\prime$ such that $\Delta_{ev^\prime,g0}$ is small).

To understand the vibronic structure of electronic resonance, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode so that hyper-Raman scattering is between this mode and the ground state. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{v'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | v' \rangle
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar} \sum_{v'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | v' \rangle \mel{v'}{Q}{0} 
				\frac{1}{\Delta_{ev', g0}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{v'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{v'} 
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
			\end{split}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $ is the electronic two photon transition moment between $a$ and $b$.\cite{McClain1977}

\autoref{ABterms_DR} shows important selection rules for the vibronic transitions.
The $A$ term is allowed whenever an electronic transition has a one- and two- photon transition moment.
For centrosymmetric molecules, an electronic transition cannot be both one- and two- photon allowed by parity, so the $A$ term will vanish.
This makes HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects.\cite{Olson2018}
For an allowed electronic transition, the brightness of a vibronic resonance is controlled by overlap of ground and excited state vibrational wavefunctions (Franck-Condon factors).
The $A$ term relies on excited state perturbations to the vibrational wavefunctions; if the excited electronic state minimally changes potential energy surface, so that ground and excited state vibrational eigenfunctions are identical, $\langle 1 | v^\prime \rangle \langle v^\prime | 0 \rangle = 0$ and $A$ vanishes.
%todo: note this is the OPPOSITE of hyper raman 
Some $B$ terms do survive in the same limit, however; for instance, $B_1$ survives for $v^\prime=1$ and $B_2$ survives for $v^\prime=0$.
In the absence of an $A$ term, the $B$ terms will dominate.
The $B_1$ and $B_2$ terms involve different types of vibronic coupling pathways.
In $B_1$, the Herzberg-Teller coupling is associated with the two-photon absorption process, whereas in $B_2$, the Herzberg-Teller coupling is associated with one-photon emission.

The signatures of vibronic coupling in the HDFG spectrum is made clear by using a simple model.
Using a two-well system described by 
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\xi)^2 \right) + \hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ (momentum) and $q$ (position) are the dimensionless conjugate coordinates of the normal mode, the A and B contributions can be evaluated using Franck-Condon and Herzberg-Teller integrals tabulated in terms of the dimensionless normal mode $q$ and offset parameter $\xi=0.5$ (\autoref{fig:doubres_spec}a).\cite{Carlson1988thesis} % DDK: there is probably a better citation than a thesis.  Carlson wasn't the first to derive these...
For simplicity, we restrict consideration to $v^\prime \in \{0,1,2\}$.
Note that $q = \sqrt{\frac{m^2\omega}{\hbar}} Q$.
We assume the electronic state is both one and two photon allowed.

@@

\autoref{fig:doubres_spec}b shows the simulated $\omega_2+\omega_3$ spectrum for this two-well system.  The spectrum selectively probes vibronic pathways from $\ket{g,0}$ to $\ket{g,1}$.
In the resonant region, the $A$ and $B$ terms contain character from all three possible transitions between $\ket{g,0}$, $\ket{e,v'}$ and $\ket{g,1}$. 
% This discussion assumes that mode symmetry allows for coupling between all of the vibronic states considered here.  ddk: there is only one mode!
Significant contributions from $\ket{e,0}$ and $\ket{e,1}$ in the $A$ term spectrum correspond well with expectations given the relative wavefunction overlap in \autoref{fig:doubres_spec}a, tabulated in \autoref{T:contributions}. 
Similarly, the $B_1$ term is dominated by coupling with the $\ket{e,1}$ state and $B_2$ is dominated by coupling with the $\ket{e,0}$ state (\autoref{T:contributions}). 
Note that contributions involving $\ket{e,2}$ are suppressed relative to the $\ket{e,0}$ and $\ket{e,1}$ contributions in the $A$ and $B$ terms.
This is expected, considering the large change in quantum number when involving $\ket{e,2}$ and the offset of the potential wells in \autoref{Hamiltonian}.

In the pre-resonance region, both $A$ and $B$ terms die off as $\sim \Delta_{e0,g0}^{-1}$.
We point this out because a persistent belief in Raman analysis is that $A$ terms disappear and $B$ terms dominate when significantly detuned from electronic resonance.
% such that $\Delta_{ev', g0} \approx \Delta_{eg}$, i.e., the resonance denominator loses $\ket{v'}$ dependence. 
The belief arises from a the approximation $\abs{\omega_{ev',g0} - 2\omega_2} \gg 2\Gamma_{ev',g0}$, which then removes $A$ through a closure relationship.\cite{Neddersen1989}
As has been discussed elsewhere, however, this approximation is in many cases invalid, especially when vibrational energies are large, like those explored here.\cite{Warshel1977, Li1990, Gong2015}
This means that it is difficult to isolate $B$ terms when $A$ terms are present.
As such, HDFG will have difficulty examining non-Condon effects ($B$ term contributions) noncentrosymmetric systems.

\begin{table}[!htbp]
	\caption{\label{T:contributions} Franck-Condon and/or Herzberg-Teller contributions to $A$, $B_1$, $B_2$ terms between $\ket{g,0}$, $\ket{g,1}$ and vibronic states $\ket{e,0}$, $\ket{e,1}$, $\ket{e,2}$ for \autoref{Hamiltonian} with an offset $\xi = 0.5$.
	Closed-form expressions for the Franck-Condon and Herzberg-Teller integrals used here have been tabulated by Carlson.\cite{Carlson1988thesis}}
	\begin{ruledtabular}
		\begin{tabular}{cccc}
			$2\omega_2$ Electronic Resonance & $\ket{e,0}$ & $\ket{e,1}$ & $\ket{e,2}$\\
			\hline  
			$A$ 
			& $\langle 1 | 0 \rangle \langle 0 | 0 \rangle = 0.31 $  
			& $\langle 1 | 1 \rangle \langle 1 | 0 \rangle = 0.27 $ 
			& $\langle 1 | 2 \rangle \langle 2 | 0 \rangle = -0.04$\\

@@

\section{Quantitative Aspects of HDFG}\label{quant}
% Despite their increasing use for dissecting intra and intermolecular interactions, CMDS techniques are usually not feasible for most laboratories, as they usually demand the use of multiple optical parametric amplifiers and/or complex acousto-optic modulators. \cite{Chen2016}
% Unlike most four wave mixing techniques, HDFG can be performed using only two input beams (i.e., two color HDFG).
% In the case of two color HDFG, only one tunable infrared optical parametric amplifier is needed, as the oscillator which pumps the infrared OPA can be used to perform the two-photon upconversion needed to generate HDFG output.
% Vibrational sum frequency generation (vSFG) is a $\chi^{(2)}$ technique whose output, when only vibrationally resonant, scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor, whose phase-matching constraint is  $\vec{k}_3 = \vec{k}_1 + \vec{k}_2$.
% HDFG is phase-matchable under $\vec{k}_4 = -\vec{k}_1 + 2\vec{k}_2$ conditions for most, if not all, vibrational resonances.
% Laboratories which use vSFG to investigate interfacial species at buried interfaces thus have the ingredients necessary to perform a HDFG experiment in a transmission geometry. \cite{Piontek2023_1}
% This would allow vSFG practitioners to perform measurements in the bulk, such as measuring steady-state vibrational spectra, free induction decay and vibrational population lifetimes through HDFG or pump-HDFG-probe.

\subsection{vSFG detection levels are adequate for HDFG}

Can vSFG experimental setups detect HDFG signals?
SFG is often consider weak compared to third order spectroscopies because it is usually surface selective,\cite{RN132, RN133} so vSFG detection setups are often quite sensitive.
However, HDFG is a weak $\chi^{(3)}$ process because it depends upon hyper-Raman cross-sections, which are commonly several orders of magnitude weaker than corresponding Raman transitions.\cite{RN515}
To evaluate the feasibility of HDFG for vSFG practitioners, we have compared the relative output of each process using nominal parameters. 
As calculated in \autoref{appendixA}, HDFG's weak $\chi^{(3)}$ value and SFG's low emission cancel somewhat, so that the output polarization of HDFG and vSFG are comparable ($\abs{P_\text{vSFG} / P_\text{HDFG}} \sim 1$).
Transient HDFG, or IR-pump-HDFG-probe, analogous to IR-pump-vSFG-probe, is therefore also feasible under similar experimental conditions, and provides vSFG practitioners a simple method to measure bulk dynamics without significant changes in an experimental setup. 
Therefore, HDFG is a feasible spectroscopy for practitioners of vSFG both in terms of experimental setup and detection feasibility.

% With the selection rules of HDFG understood for vibrational spectroscopy, and with HDFG generating a large enough output polarization, it becomes possible to obtain quantitative information from its spectra.
% Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
% Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re(\chi^{(3)})$ and $\Im(\chi^{(3)})$.\cite{Levenson1974_1, Levenson1974_2}
% In the method introduced by Levenson and Bloembergen (Bloembergen interferometry experiment), an internal standard interferes with the resonant lineshape.
% Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
% Resonant lineshapes are also complicated by amplitude level interference between the sample,  substrate and/or sample cell windows, which must be accounted for to obtain quantitatively correct $\chi^{(3)}$ values. \cite{RN362, RN418}
% Most quantitative methods in an $n^{th}$ order CMDS experiment are used to measure $\chi^{(n)}$ values to compare the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
% The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.

\subsection{Extraction of hyper-Raman cross-sections}

% DDK: make sure statements are correct here
Compared to $\vec{\mu}$ and $\alpha_{ij}$, it is difficult to calculate $\beta_{ijk}$ values from spontaneous hyper-Raman experiments.\cite{Kelley2010}
Only a few experimental determinations of $\beta_{ijk}$ for vibrational modes have been performed.\cite{Xu1997, Shoute2005, Kelley2010}
Methods reported in the literature to measure absolute hyper-Raman polarizabilities depend upon internal standards such as hyperpolarizabilities of dissolved samples or two-photon absorption cross sections, also difficult to measure. \cite{Okuno2020}  % DDK:  correct reference?
Since the Bloembergen interferometry experiment only relies on the third order susceptibility for measured species (e.g., benzene or CaF$_2$),\cite{Levenson1974_2} it is possible to use quantitative four wave mixing spectroscopy to calculate $\beta_{ijk}$ values.
It is thus useful to investigate how a treatment of orientational averaging can extract $\beta_{ijk}$ from the HDFG $\chi^{(3)}_{IJKL}$ expression.

Using \autoref{eq:nfgamma} and \autoref{sivebeta}, we can write
\begin{equation}\label{chi3}
		\chi^{(3)}_{IJKL} = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle \rho_{gg}\\
\end{equation}
For simplicity, we take $\rho_{gg} = 1$.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Briefly, a tensor in the molecular frame, $A_{ijkl}$, is transformed into an element of the same tensor in the laboratory frame, $\langle A \rangle_{IJKL}$, through $\langle A \rangle_{IJKL}$ = $\theta^{ijkl}_{IJKL} A_{ijkl}$, where summation over repeated indices is implied and $\theta$ is the transformation operator. \cite{McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Willetts1992, Kauranen1996}
Substituting in \autoref{SIVEselection}, we see
% DDK: why does it matter that chi is imaginary?
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta}{\partial Q_v} {\frac{\partial \mu}{\partial Q_v}} \right\rangle _{IJKL} = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} \chi^{(3)}_{IJKL}
\end{equation}
$\chi^{3}$ itself can be extracted with use of an internal standard.\cite{Levenson1974_1, Levenson1974_2}
Since $\abs{\partial \mu / \partial Q}$ values can be extracted from FT-IR spectra,\cite{RN459} HDFG can give quantitative information on the magnitude of $\beta_{ijk}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Christie1971, Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values.
